Formalizing the Kolmogorov extension Theorem in Lean

### 2023-09-12 Manuscript/proof_outline.tex
minor rewrite
add leanlinem command to mark mathlib lemmas

\usepackage{minted}
\setminted{encoding=utf-8}
\usepackage{fontspec}
%\setmainfont{dejavu-serif}
\setmonofont[Scale=0.833]{dejavu-sans}
%\setmainfont{FreeSerif}
%\setmonofont{FreeMono}
\usepackage{xcolor, graphics}
\definecolor{mygray}{rgb}{0.97,0.97,0.97}
\newcommand{\leanline}{\mintinline[breaklines, breakafter=., breaklines]{Lean}}
%\usepackage{refcheck}

@@

\todo[inline]{\leanline{∀ j : J, α j} or \leanline{(j : J) -> α j} ?}

@@

    \item \sloppy In \leanline{Lean}, we have the \leanline{class PseudoEMetricSpace (α : Type u) extends EDist}, where
      \leanline{EDist} comes with an extended distance \leanline{edist: α → α → ℝ≥0∞}

@@

      \leanline{PseudoEMetricSpace} consists of the components,

@@

      uniformity; see \leanline{UniformSpace.pseudoMetrizableSpace},

@@

    \item \sloppy For topological spaces, we have \leanline{class TopologicalSpace(α : Type u) : Type u}, which comes with

@@

      a uniform space is a \leanline{class UniformSpace(α : Type u)

@@

      uniformity is defined in \leanline{UniformSpace.Core}), which

@@

\begin{remark}[Generated topology]
  \begin{enumerate}
  \item The intersection of any number of topologies is again a
    topology. For this reason, if $\mathcal H \subseteq 2^\alpha$, we
    define the topology $\mathcal O := \bigcap_{\mathcal F \supseteq
      \mathcal H \text{ topology}} \mathcal F$; see
    \leanline{TopologicalSpace.generateFrom}. This is called the
    topology generated by $\mathcal H$. If $\mathcal H$ is closed
    under finite intersections, we call $\mathcal H$ a basis for
    $\mathcal O$; see \leanline{TopologicalSpace.IsTopologicalBasis}.
  \item \sloppy We call the topology (generated from an extended
    pseudo-metric) separable (see
    \leanline{TopologicalSpace.SeparableSpace}) if there is a
    countable $s \subseteq \iota$ such that $\inf\{r(x,y) : y \in s\}
    = 0$ for all $x\in\alpha$. (More generally, a
    \leanline{TopologicalSpace α} is a \leanline{SeparableSpace} iff
    \leanline{∃ s, Set.Countable s ∧ Dense s}, where the latter means
    \leanline{∀ (x : α), x ∈ closure s}, i.e.\ the space equals its
    own closure.) \\ If there is a countable basis of the topology, it
    is separable; see
    \leanline{TopologicalSpace.SecondCountableTopology.to_separableSpace}.
  \end{enumerate}
\end{remark}

@@

\begin{remark}[Measur(abl)e spaces in \leanline{mathlib}]
  \begin{enumerate}
  \item \sloppy The \leanline{class MeasurableSpace α} is very similar
    to a topological space in \leanline{mathlib}, since it comes with
    \leanline{MeasurableSet' : Set α → Prop},
    \leanline{measurableSet_empty : MeasurableSet' ∅},
    \leanline{measurableSet_compl : (s : Set α) → MeasurableSet' s → MeasurableSet' sᶜ}, and \leanline{measurableSet_iUnion},
    i.e.\ properties (i)-(iii) from Definition~\ref{def:sigma}.1.
  \item We are using outer measures in our construction. In
    \leanline{mathlib}, this is \leanline{structure MeasureTheory.OuterMeasure α} which is a function
    \leanline{measureOf : Set α → ℝ≥0∞} such that \leanline{measureOf ∅ = 0} (i.e.\ the empty set has measure~$0$),
    \leanline{∀ {s₁ s₂ : Set α}, s₁ ⊆ s₂ → measureOf s₁ ≤ measureOf s₂}
    (i.e.\ monotonicity) and \leanline{∀ (s : ℕ → Set α), measureOf (⋃ (i : ℕ), s i) ≤ ∑' (i : ℕ), measureOf (s i)} (which we call
    $\sigma$-subadditivity). \\ Note that an outer measure is defined
    on $2^\alpha$, whereas a measure is only defined on a subset (the
    $\sigma$-algebra of measurable sets).
  \item A measure defined on a type \leanline{α} on which we have a \leanline{MeasurableSpace α} is denoted by the type \leanline{MeasureTheory.Measure α}
    %More precisely\todo{I don't understand why all these details are here. We never use MeasureSpace}: A
    %\leanline{MeasureSpace α} extends a \leanline{MeasurableSpace α}
    %by a \leanline{Measure α} (which is usually called
    %\leanline{volume}). For the measure, we have: \leanline{structure MeasureTheory.Measure α [MeasurableSpace α] extends MeasureTheory.OuterMeasure}.
    %Here $\sigma$-additivity of a
    %\leanline{Measure α} reads
%\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
%  m_iUnion : ∀ ⦃f : ℕ → Set α⦄,
%    (∀ (i : ℕ), MeasurableSet (f i)) →
%    Pairwise (Disjoint on f) →
%    ↑toOuterMeasure (⋃ (i : ℕ), f i) =
%    ∑' (i : ℕ), ↑toOuterMeasure (f i)
%\end{minted}    
  \end{enumerate}
\end{remark}

@@

\begin{remark}[Generated $\sigma$-algebra, image measure]
  We will frequently need two basic results: 
  \begin{enumerate}
  \item The intersection of any number of $\sigma$-algebras is again a
    $\sigma$-algebra. For this reason, if $\mathcal H \subseteq
    2^\alpha$, we define the $\sigma$-algebra $\sigma(\mathcal H) :=
    \bigcap_{\mathcal F \supseteq \mathcal H \text{ $\sigma$-algebra}}
    \mathcal F$. This is called the $\sigma$-algebra generated from
    $\mathcal H$; see \leanline{MeasurableSpace.generateFrom}.  \\ In
    particular, if $\mathcal O$ is a topology on $\alpha$, we call
    $\mathcal B := \sigma(\mathcal O)$ the Borel $\sigma$-algebra on
    $\alpha$. In \leanline{mathlib}, this is
\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  def borel (α : Type u) [TopologicalSpace α] :
    MeasurableSpace α :=
    generateFrom { s : Set α | IsOpen s }
\end{minted}
    In addition, \leanline{mathlib} provides a similar notion, which is
\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  class OpensMeasurableSpace (α : Type*)
    [TopologicalSpace α] [h : MeasurableSpace α] :
    Prop where borel_le : borel α ≤ h
\end{minted}
    Here, all open sets are measurable, so the $\sigma$-algebra defining
    \leanline{h} might be larger than the Borel $\sigma$-algebra. We will
    need both notions in our proofs.
  \item Let $(\alpha, \mathcal F)$ and $(\beta, \mathcal G)$ be
    measurable spaces. Some $f : \alpha \to \beta$ is called
    measurable (with respect to $\mathcal F$ and $\mathcal G$) if
    $f^{-1} \mathcal G \subseteq \mathcal F$. In \leanline{mathlib},
    see \leanline{Measurable}.
    %and note that this notion usually comes
    %with some \leanline{m : Measure α} and $f$ is almost everywhere
    %measurable (\leanline{AEMeasurable}) if it coincices with some
    %\leanline{Measurable g}, \leanline{m}-almost everywhere\todo{do we need aemeasurable? could it be omitted for simplicity?}.
    \\ In this
    case, if $\mu$ is a measure on $\mathcal F$, the measure $\nu:
    \mathcal G \to [0, \infty], \nu(B) := \mu (f^{-1}(B))$ is called
    the image (or push-forward) measure of $\mu$ under $f$. We write
    $\nu := f_\ast \mu$; see \leanline{MeasureTheory.Measure.map}.
  \end{enumerate}
\end{remark}

We now come to our contribution%, which we will PR to mathlib
: the extension theorem is on finite measures on product spaces\todo{I don't understand this sentence}. In
particular, the product is taken over a set of arbitrary
cardinality. The next definition covers the important concept of a
projective family of measures. In short, we define measures on
any finite subset of indices in a consistent way. Finite
sets (of some type \leanline{ι}) are formalized with \leanline{Finset ι}. % Such a \leanline{Finset} is defined as a \leanline{Multiset} (a
%quotient type of the \leanline{List}-type with respect to permutations
%of the list) together with a proof that it contains no duplicates\todo{remove this implementation detail?}.

@@

Let us remark that (i) \leanline{Lean} indicates a coercion by
\leanline{↑}, which in this case is from \leanline{Finset} to
\leanline{Set} and (ii) we have \leanline{PairwiseDisjoint (s : Set ι) (f : ι → α)}
iff the images of any distinct elements of \leanline{ι}
under \leanline{f} are different. (Hence, if \leanline{f = id}, the
usual definition of pairwise disjoint sets unfolds.)

\noindent
We do not show the formalization of rings here.
%In our formalization, we also use the notion of a field $\mathcal H$,
%which is a ring with $\alpha \in
%\mathcal H$.
Since we are dealing with product spaces, we will use cylinders\todo{R: rewrite this}, which
are defined as follows:

@@

\noindent
The existence-statement of the above lemma actually gives rise to a
definition. Here, we use \leanline{Exists.choose} in order to extract
an element from an $\exists$-statement. In addition, we do not allow
$\emptyset \in \mathcal K$ without loss of generality\todo{that exclusion was useful for something. Find out what that is}:

@@

Here, recall that \leanline{Fin n} is the subtype of \leanline{Nat}
consisting of all numbers \leanline{<n}. In addition,
\leanline{ordered_mem' (hJ : ↑J ⊆ C) (n : Fin (card J)) : ↑(ordered J)
  n ∈ C} gives a proof that the \leanline{n}th element of \leanline{J}
is in \leanline{C}, \leanline{finsetLT (J : Finset (Set α)) (n : Fin (card J)) : Finset (Set α)}
is the set consisting of the
\leanline{n} sets in \leanline{J} numbered $0,...,n-1$, and
\leanline{finsetLT_subset'} gives a proof that \leanline{↑(finsetLT J n) ⊆ C}.
So, in terms of Lemma~\ref{l2}, this gives, for
$n=1,...,m$, a construction for $\mathcal K_n \subseteq_f \mathcal H$.

@@

\noindent
For additive contents\todo{"content" not defined?}, we need two definitions. \leanline{AddContent}
defines a set-function \leanline{tuFun} on \leanline{Set α}, whereas
\leanline{extendContent} comes with \leanline{m : ∀ s : Set α, s ∈ C → ℝ≥0∞}.

@@

\sloppy Let us now formulate the classical results by
Carathéodory. The first leads to the definition of the measurable
space \leanline{OuterMeasure.caratheodory}, covered in {\tt
  mathlib}. See e.g.\ \cite[Theorem 2.1]{Kallenberg2020}

@@

\noindent
\sloppy The second result states that for an outer measure induced by
an additive content on a semi-ring, we have $\sigma(\mathcal H)
\subseteq \mathcal F$. In particular, we then have defined a measure
on $\sigma(\mathcal H)$. This result is not yet covered in {\tt
  mathlib}. However, for product spaces, in
\leanline{MeasureTheory.Constructions.Pi}, there is
\leanline{pi_caratheodory}, which shows that $\sigma(\mathcal H)
\subseteq \mathcal F$ in a special case\todo{which special case?}. In addition, there is
\leanline{pi_pi_aux}, which shows that $\mu$ extends $m$ on $\mathcal
H$ in the same special case. See also \cite[Theorem
  2.5]{Kallenberg2020}.

@@

\noindent
Here, \leanline{MeasureTheory.OuterMeasure.caratheodory}\todo{used in dot notation... but we don't want to explain this} is already
implemented and gives the assertion that $\mathcal F$ from
Theorem~\ref{T:cara} is a $\sigma$-algebra. Note that
\leanline{@Measure} is the same as \leanline{Measure}, but all
variables have to be given explicitely. The second argument is a proof
that measurable sets form a $\sigma$-algebra which cannot be inferred
by the type class system of \leanline{Lean}.

@@

In our formalization, we are less general and use a direct argument
for a \leanline{kol_content}. More precisely, we show that a
\leanline{kolContent}, which is based on a projective family which is
inner regular with respect to compact and closed sets is
$\sigma$-additive.

@@

The next result is already implemented in \leanline{mathlib}: Use
\leanline{Measure.InnerRegular.of_pseudoEMetricSpace} in order to show
that every open set is inner regular with respect to closed sets, and
\leanline{Measure.InnerRegular.measurableSet_of_open} to show that
every measurable set is inner regular with respect to closed sets.
The claimed outer reularity is part of
\leanline{MeasureTheory.Measure.InnerRegular.weaklyRegular_of_finite}.

@@

In the next lemma, we will need that a closed subset of a compact set
is compact. (See \leanline{isCompact_of_isClosed_subset}.) We describe
the formalization of the next lemma only in conjunction with
Lemma~\ref{L:relcoPol}.

@@

We apply this result by using that 1.\ is satisfied for complete and
separable extended pseudo-metric spaces. Here, for the generalization
to extended pseudo-metric spaces, we use that any subset of such a
space is compact iff it is complete and totally bounded (see
\leanline{isCompact_iff_totallyBounded_isComplete} in {\tt
  mathlib}). Since closed subsets of a complete space are complete,
the closure of a totally bounded set is (still totally bounded, hence)
compact. (This fact is used in the proof below.)

@@

{\em Using semi-rings in measure theory:} Frequently, the construction
of measures on some $\sigma$-algebra $\mathcal F$ (on some space
$\alpha$) uses outer measures, which can be defined on $2^\alpha$ (the
set of all subsets of $\alpha$). We provide a general framework using
Carathéodory's extension theorem, which states an extension of a
set-function on a semi-ring by an outer measure, not previously
implemented in mathlib. This could also be used to re-define Stieltjes
measures (in particular Lebesgue measure) on $\mathbb R$, as well as
product measures (on finite products). More precisely,
e.g.\ \leanline{StieltjesFunction.outer_Ioc} and
\leanline{MeasureTheory.constructions.pi.pi_pi_aux} are specific
examples of the more general statement, that the outer measure extends
the set function a the semi-ring. (In the first case, it would be the
semi-ring of half-open intervals, in the second case, it would be the
semi-ring of cylinder sets on a product space.)

>>>>

\usepackage{minted}
\setminted{encoding=utf-8}
\usepackage{fontspec}
%\setmainfont{dejavu-serif}
\setmonofont[Scale=0.833]{dejavu-sans}
%\setmainfont{FreeSerif}
%\setmonofont{FreeMono}
\usepackage{xcolor, graphics}
\definecolor{mygray}{rgb}{0.97,0.97,0.97}
\newcommand{\leanline}{\mintinline[breaklines, breakafter=., breaklines]{Lean}}
\newcommand{\leanlinem}[1]{\mintinline[breaklines, breakafter=., breaklines]{Lean}{#1}\textsuperscript{\textcolor{gray}{\tiny(m)}}}
%\usepackage{refcheck}

@@

\todo[inline]{\leanline{∀ j : J, α j} or \leanlinem{(j : J) -> α j} ?}

\todo[inline]{explain leanlinem}

@@

    \item \sloppy In \leanline{Lean}, we have the \leanlinem{class PseudoEMetricSpace (α : Type u) extends EDist}, where
      \leanlinem{EDist} comes with an extended distance \leanline{edist: α → α → ℝ≥0∞}

@@

      \leanlinem{PseudoEMetricSpace} consists of the components,

@@

      uniformity; see \leanlinem{UniformSpace.pseudoMetrizableSpace},

@@

    \item \sloppy For topological spaces, we have \leanlinem{class TopologicalSpace(α : Type u) : Type u}, which comes with

@@

      a uniform space is a \leanlinem{class UniformSpace(α : Type u)

@@

      uniformity is defined in \leanlinem{UniformSpace.Core}), which

@@

\begin{remark}[Generated topology]
  \begin{enumerate}
  \item The intersection of any number of topologies is again a
    topology. For this reason, if $\mathcal H \subseteq 2^\alpha$, we
    define the topology $\mathcal O := \bigcap_{\mathcal F \supseteq
      \mathcal H \text{ topology}} \mathcal F$; see
    \leanlinem{TopologicalSpace.generateFrom}. This is called the
    topology generated by $\mathcal H$. If $\mathcal H$ is closed
    under finite intersections, we call $\mathcal H$ a basis for
    $\mathcal O$; see \leanlinem{TopologicalSpace.IsTopologicalBasis}.
  \item \sloppy We call the topology (generated from an extended
    pseudo-metric) separable (see
    \leanlinem{TopologicalSpace.SeparableSpace}) if there is a
    countable $s \subseteq \iota$ such that $\inf\{r(x,y) : y \in s\}
    = 0$ for all $x\in\alpha$. (More generally, a
    \leanline{TopologicalSpace α} is a \leanlinem{SeparableSpace} iff
    \leanline{∃ s, Set.Countable s ∧ Dense s}, where the latter means
    \leanline{∀ (x : α), x ∈ closure s}, i.e.\ the space equals its
    own closure.) \\ If there is a countable basis of the topology, it
    is separable; see
    \leanlinem{TopologicalSpace.SecondCountableTopology.to_separableSpace}.
  \end{enumerate}
\end{remark}

@@

\begin{remark}[Measur(abl)e spaces in \leanline{mathlib}]
  \begin{enumerate}
  \item \sloppy The \leanlinem{class MeasurableSpace α} is very similar
    to a topological space in \leanline{mathlib}, since it comes with
    \leanline{MeasurableSet' : Set α → Prop},
    \leanline{measurableSet_empty : MeasurableSet' ∅},
    \leanline{measurableSet_compl : (s : Set α) → MeasurableSet' s → MeasurableSet' sᶜ}, and \leanline{measurableSet_iUnion},
    i.e.\ properties (i)-(iii) from Definition~\ref{def:sigma}.1.
  \item We are using outer measures in our construction. In
    \leanline{mathlib}, this is \leanlinem{structure MeasureTheory.OuterMeasure α} which is a function
    \leanline{measureOf : Set α → ℝ≥0∞} such that \leanline{measureOf ∅ = 0} (i.e.\ the empty set has measure~$0$),
    \leanline{∀ {s₁ s₂ : Set α}, s₁ ⊆ s₂ → measureOf s₁ ≤ measureOf s₂}
    (i.e.\ monotonicity) and \leanline{∀ (s : ℕ → Set α), measureOf (⋃ (i : ℕ), s i) ≤ ∑' (i : ℕ), measureOf (s i)} (which we call
    $\sigma$-subadditivity). \\ Note that an outer measure is defined
    on $2^\alpha$, whereas a measure is only defined on a subset (the
    $\sigma$-algebra of measurable sets).
  \item A measure defined on a type \leanline{α} on which we have a \leanline{MeasurableSpace α} is denoted by the type \leanlinem{MeasureTheory.Measure α}
    %More precisely\todo{I don't understand why all these details are here. We never use MeasureSpace}: A
    %\leanline{MeasureSpace α} extends a \leanline{MeasurableSpace α}
    %by a \leanline{Measure α} (which is usually called
    %\leanline{volume}). For the measure, we have: \leanline{structure MeasureTheory.Measure α [MeasurableSpace α] extends MeasureTheory.OuterMeasure}.
    %Here $\sigma$-additivity of a
    %\leanline{Measure α} reads
%\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
%  m_iUnion : ∀ ⦃f : ℕ → Set α⦄,
%    (∀ (i : ℕ), MeasurableSet (f i)) →
%    Pairwise (Disjoint on f) →
%    ↑toOuterMeasure (⋃ (i : ℕ), f i) =
%    ∑' (i : ℕ), ↑toOuterMeasure (f i)
%\end{minted}    
  \end{enumerate}
\end{remark}

@@

\begin{remark}[Generated $\sigma$-algebra, image measure]
  We will frequently need two basic results: 
  \begin{enumerate}
  \item The intersection of any number of $\sigma$-algebras is again a
    $\sigma$-algebra. For this reason, if $\mathcal H \subseteq
    2^\alpha$, we define the $\sigma$-algebra $\sigma(\mathcal H) :=
    \bigcap_{\mathcal F \supseteq \mathcal H \text{ $\sigma$-algebra}}
    \mathcal F$. This is called the $\sigma$-algebra generated from
    $\mathcal H$; see \leanlinem{MeasurableSpace.generateFrom}.  \\ In
    particular, if $\mathcal O$ is a topology on $\alpha$, we call
    $\mathcal B := \sigma(\mathcal O)$ the Borel $\sigma$-algebra on
    $\alpha$. In \leanline{mathlib}, this is
\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  def borel (α : Type u) [TopologicalSpace α] :
    MeasurableSpace α :=
    generateFrom { s : Set α | IsOpen s }
\end{minted}
    In addition, \leanline{mathlib} provides a similar notion, which is
\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  class OpensMeasurableSpace (α : Type*)
    [TopologicalSpace α] [h : MeasurableSpace α] :
    Prop where borel_le : borel α ≤ h
\end{minted}
    Here, all open sets are measurable, so the $\sigma$-algebra defining
    \leanline{h} might be larger than the Borel $\sigma$-algebra. We will
    need both notions in our proofs.
  \item Let $(\alpha, \mathcal F)$ and $(\beta, \mathcal G)$ be
    measurable spaces. Some $f : \alpha \to \beta$ is called
    measurable (with respect to $\mathcal F$ and $\mathcal G$) if
    $f^{-1} \mathcal G \subseteq \mathcal F$. In \leanline{mathlib},
    see \leanlinem{Measurable}.
    %and note that this notion usually comes
    %with some \leanline{m : Measure α} and $f$ is almost everywhere
    %measurable (\leanline{AEMeasurable}) if it coincices with some
    %\leanline{Measurable g}, \leanline{m}-almost everywhere\todo{do we need aemeasurable? could it be omitted for simplicity?}.
    \\ In this
    case, if $\mu$ is a measure on $\mathcal F$, the measure $\nu:
    \mathcal G \to [0, \infty], \nu(B) := \mu (f^{-1}(B))$ is called
    the image (or push-forward) measure of $\mu$ under $f$. We write
    $\nu := f_\ast \mu$; see \leanlinem{MeasureTheory.Measure.map}.
  \end{enumerate}
\end{remark}

We now come to our contribution%, which we will PR to mathlib
: the extension theorem is on finite measures on product spaces\todo{I don't understand this sentence}. In
particular, the product is taken over a set of arbitrary
cardinality. The next definition covers the important concept of a
projective family of measures. In short, we define measures on
any finite subset of indices in a consistent way. Finite
sets (of some type \leanline{ι}) are formalized with \leanlinem{Finset ι}. % Such a \leanline{Finset} is defined as a \leanline{Multiset} (a
%quotient type of the \leanline{List}-type with respect to permutations
%of the list) together with a proof that it contains no duplicates\todo{remove this implementation detail?}.

@@

Let us remark that (i) \leanline{Lean} indicates a coercion by
\leanline{↑}, which in this case is from \leanline{Finset} to
\leanline{Set} and (ii) we have \leanlinem{PairwiseDisjoint (s : Set ι) (f : ι → α)}
iff the images of any distinct elements of \leanline{ι}
under \leanline{f} are different. (Hence, if \leanline{f = id}, the
usual definition of pairwise disjoint sets unfolds.)

\noindent
We do not show the formalization of rings here.
%In our formalization, we also use the notion of a field $\mathcal H$,
%which is a ring with $\alpha \in
%\mathcal H$.
The import ring of sets in our formalization is the ring of measurable cylinders on a product space, which
are defined as follows:

@@

\noindent
The existence-statement of the above lemma actually gives rise to a
definition. Here, we use \leanlinem{Exists.choose} in order to extract
an element from an $\exists$-statement. In addition, we do not allow
$\emptyset \in \mathcal K$ without loss of generality\todo{that exclusion was useful for something. Find out what that is}:

@@

Here, recall that \leanlinem{Fin n} is the subtype of \leanline{Nat}
consisting of all numbers \leanline{<n}. In addition,
\leanline{ordered_mem' (hJ : ↑J ⊆ C) (n : Fin (card J)) : ↑(ordered J)
  n ∈ C} gives a proof that the \leanline{n}th element of \leanline{J}
is in \leanline{C}, \leanline{finsetLT (J : Finset (Set α)) (n : Fin (card J)) : Finset (Set α)}
is the set consisting of the
\leanline{n} sets in \leanline{J} numbered $0,...,n-1$, and
\leanline{finsetLT_subset'} gives a proof that \leanline{↑(finsetLT J n) ⊆ C}.
So, in terms of Lemma~\ref{l2}, this gives, for
$n=1,...,m$, a construction for $\mathcal K_n \subseteq_f \mathcal H$.

@@

\noindent
For additive contents\todo{"content" not defined?}, we need two definitions. \leanline{AddContent}
defines a set-function \leanline{toFun} on \leanline{Set α}, whereas
\leanline{extendContent} comes with \leanline{m : ∀ s : Set α, s ∈ C → ℝ≥0∞}.

@@

\sloppy Let us now formulate the classical results by
Carathéodory. The first leads to the definition of the measurable
space \leanlinem{OuterMeasure.caratheodory}, covered in {\tt
  mathlib}. See e.g.\ \cite[Theorem 2.1]{Kallenberg2020}

@@

\noindent
\sloppy The second result states that for an outer measure induced by
an additive content on a semi-ring, we have $\sigma(\mathcal H)
\subseteq \mathcal F$. In particular, we then have defined a measure
on $\sigma(\mathcal H)$. This result is not yet covered in {\tt
  mathlib}. However, for product spaces, in
\leanlinem{MeasureTheory.Constructions.Pi}, there is
\leanlinem{pi_caratheodory}, which shows that $\sigma(\mathcal H)
\subseteq \mathcal F$ in a special case\todo{which special case?}. In addition, there is
\leanlinem{pi_pi_aux}, which shows that $\mu$ extends $m$ on $\mathcal
H$ in the same special case. See also \cite[Theorem
  2.5]{Kallenberg2020}.

@@

\noindent
Here, \leanlinem{MeasureTheory.OuterMeasure.caratheodory}\todo{used in dot notation... but we don't want to explain this} is already
implemented and gives the assertion that $\mathcal F$ from
Theorem~\ref{T:cara} is a $\sigma$-algebra. Note that
\leanline{@Measure} is the same as \leanline{Measure}, but all
variables have to be given explicitely. The second argument is a proof
that measurable sets form a $\sigma$-algebra which cannot be inferred
by the type class system of \leanline{Lean}.

@@

In our formalization, we are less general and use a direct argument
for a \leanline{kolContent}. More precisely, we show that a
\leanline{kolContent}, which is based on a projective family which is
inner regular with respect to compact and closed sets is
$\sigma$-additive.

@@

The next result is already implemented in \leanline{mathlib}: Use
\leanlinem{Measure.InnerRegular.of_pseudoEMetricSpace} in order to show
that every open set is inner regular with respect to closed sets, and
\leanlinem{Measure.InnerRegular.measurableSet_of_open} to show that
every measurable set is inner regular with respect to closed sets.
The claimed outer reularity is part of
\leanlinem{MeasureTheory.Measure.InnerRegular.weaklyRegular_of_finite}.

@@

In the next lemma, we will need that a closed subset of a compact set
is compact. (See \leanlinem{isCompact_of_isClosed_subset}.) We describe
the formalization of the next lemma only in conjunction with
Lemma~\ref{L:relcoPol}.

@@

We apply this result by using that 1.\ is satisfied for complete and
separable extended pseudo-metric spaces. Here, for the generalization
to extended pseudo-metric spaces, we use that any subset of such a
space is compact iff it is complete and totally bounded (see
\leanlinem{isCompact_iff_totallyBounded_isComplete} in {\tt
  mathlib}). Since closed subsets of a complete space are complete,
the closure of a totally bounded set is (still totally bounded, hence)
compact. (This fact is used in the proof below.)

@@

{\em Using semi-rings in measure theory:} Frequently, the construction
of measures on some $\sigma$-algebra $\mathcal F$ (on some space
$\alpha$) uses outer measures, which can be defined on $2^\alpha$ (the
set of all subsets of $\alpha$). We provide a general framework using
Carathéodory's extension theorem, which states an extension of a
set-function on a semi-ring by an outer measure, not previously
implemented in mathlib. This could also be used to re-define Stieltjes
measures (in particular Lebesgue measure) on $\mathbb R$, as well as
product measures (on finite products). More precisely,
e.g.\ \leanlinem{StieltjesFunction.outer_Ioc} and
\leanlinem{MeasureTheory.constructions.pi.pi_pi_aux} are specific
examples of the more general statement, that the outer measure extends
the set function a the semi-ring. (In the first case, it would be the
semi-ring of half-open intervals, in the second case, it would be the
semi-ring of cylinder sets on a product space.)
